Co-moving stars in \textsl{Gaia DR1}: a catalog of genuine comoving star pairs from low-resolution spectroscopic follow-up

### 2017-05-31 paper/comoving-followup.tex
first figure
more text and shit

\documentclass[modern, letterpaper]{aastex61}

@@

\include{gitstuff}
\include{preamble}

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}1}
\newcommand{\tgas}{\acronym{TGAS}}

@@

\subsection{Observations and Data reduction}\label{sec:reduction}

@@

For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, the 2D spectra are extracted to 1D using an
empirically-estimated line spread function (LSF): at each of 16 pixel rows
between indices 750 and 850 (i.e. around the effective center of the dispersion
axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept at
the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from each
pixel row using \texttt{L-BFGS-B} optimization (\citealt{Zhu:1994}), implemented
in the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source image, the LSF amplitude and centroid, and
background slope and intercept at line centroid are then fit using the same
procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

\package{celerite} (\citealt{Foreman-Mackey:2017})

>>>>

\documentclass[modern, letterpaper]{aastex61}
% \documentclass[twocolumn]{aastex61}

@@

\include{gitstuff}
\include{preamble}
\graphicspath{{figures/}}

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}

@@

\todo{color-magnitude diagram of all stars observed, color our targets different
from the calibration targets}

\subsection{Observations and data reduction}\label{sec:reduction}

@@

For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, the 2D spectra are extracted to 1D using an
empirically-estimated line spread function (LSF): at each of 16 pixel rows
between indices 750 and 850 (i.e. around the effective center of the dispersion
axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept at
the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from each
pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented
in the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source image, the LSF amplitude and centroid, and
background slope and intercept at line centroid are then fit using the same
procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

We start the wavelength calibration by interactively identifying the 12
strongest, least blended emission lines in one of the extracted 1D comparison
lamp spectra by comparing to \todo{NIST??}.
Once a rough mapping from pixel to wavelength is determined, we then fit for the
individual line centroids in each comparison lamp spectrum by extracting small,
8 pixel-wide segments of spectrum around each identified emission line and
fitting a model for the line profile and background to each individually.
We use a Gaussian profile for each emission line with parameters: amplitude,
$A$; centroid, $x_0$; and root-variance, $\sigma_G$.
We found that nearby blended or low-amplitude lines (i.e. part of the
``background'') sometimes bias the line centroid determination when using a
simple background model.
We therefore model the background flux around each comparison lamp emission line
using a constant offset + Gaussian process (GP) with a Mat\'ern 3/2 kernel
parametrized by amplitude parameter $\sigma_{3/2}$ and correlation scale
$\rho_{3/2}$.
We use \package{celerite} (\citealt{Foreman-Mackey:2017}) to evaluate the full
likelihood, and maximize the (log-)likelihood again using \lbfgsb\ optimization.
We optimize over the logarithm of any scale parameters (i.e. $A$, $\sigma_G$,
$\sigma_{3/2}$, and $\rho_{3/2}$).
Once we determine the pixel centroids for all identified emission lines for a
given night's comparison lamp spectrum, we then fit an interpolating function to the pixel-centroid, wavelength values for each emission line.

Typically, high-order standard polynomials or Chebyshev polynomials are used to
fit for the wavelength dispersion function.
We found that using such functions can lead to vastly incorrect wavelength
values near either end of the pixel grid where the polynomial function is
un- or weakly-constrained.
We instead use a combined linear polynomial + GP to model the wavelength
dispersion.
We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
scale parameter to log-space before maximizing the log-likelihood.
\figurename~\ref{fig:wavelength-GP} shows an example of one such fit with the
best-fit linear polynomial subtracted from the (pixel-centroid, wavelength)
pairs (black points) and fit curve (darker line).
The envelope around the best-fit curve (shaded region) shows the \todo{1$\sigma$
blah blah GP...}
When the GP variance is large, the pixel-to-wavelength mapping is uncertain; we
store this variance and later use it to mask regions of the spectra with
uncertain wavelength values.
This wavelength calibration procedure produces a nightly model for the nonlinear
dispersion of wavelength as a function of pixel value for each extracted 1D
spectrum.
\figurename~\ref{fig:sample-spectra} shows a random sample of 10 source spectra
of comoving star pair members normalized to 1 at $\lambda = 5500~{\rm \AA}$ and
shifted by index for visualization.
From the appearances of these 10 alone, it seems there is a mix of metallicities
and spectral types.
Note that the spectra are not flux-calibrated, as we only care about measuring
radial velocities from this sample.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{wavelength_gp.pdf}
  \end{center}
  \caption{%
    \todo{Stuff}
    \label{fig:wavelength-GP}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{sample_spectra.pdf}
  \end{center}
  \caption{%
    \todo{Stuff}
    \label{fig:sample-spectra}}
\end{figure}

TODO: sky-line corrections to the spectra
However, because of flexure in the spectrograph, each observation will have an
additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm \AA}$,
corresponding to a significant velocity offset at ${\rm H}\alpha$ of
$\approx 20$--$100~\kms$.

TODO: line measurements, model, gaussian process
